---
title: "Mini-Project 2"
author: "Leigh Tanji & Samantha Bryce"
date: "October 31, 2017"
output: html_document
---


## Loading the data

```{r, include=FALSE}
load("house_elections.rda")
load("candidates.rda")
load("committees.rda")
load("contributions.rda")
```

```{r}
library(dplyr)
library(ggplot2)
```

```{r}
tbl_df(house_elections)
```

#How many incumbents won their elections? 
Saving this information for both parties as a dataframe.
```{r}
parties <- c("D", "R")

win_incumbs<- function(party_arg) {
house_elections %>%
  filter(incumbent == "TRUE" & party == party_arg) %>%
  group_by(state) %>%
  summarise(winners = sum(ge_winner == "W"))
}

wins <- lapply(parties, win_incumbs)

names(wins) <- c("win_dem", "win_rep")
list2env(wins, envir = .GlobalEnv)
```

#How many candidates were running for each party?
```{r}
total_candidates <- function(party_arg) {
  house_elections %>%
  filter(party == party_arg) %>%
  group_by(state) %>%
  summarise(N = n()) %>%
  data.frame()
}

total <- lapply(parties, total_candidates)

names(total) <- c("total_dem", "total_rep")
list2env(total, envir = .GlobalEnv)

tbl_df(total_dem)
tbl_df(total_rep)
```

Checking the 'total_incumbs' numbers from 'incumbs'. Should be 794 and 925. 
```{r}
total1 <- total_dem %>%
  summarise(SUM = sum(N))
total1

total2 <- total_rep %>%
  summarise(SUM = sum(N))
total2

total3 <- function(party_arg) {
  house_elections %>%
  filter(party == party_arg) %>%
  group_by(state) %>%
  select(state, party)
}

lapply(parties, total3)
```

Checking the 'winners_incumb' numbers from 'incumbs'. Should be 152 and 199.
```{r}
winners_incumbs1 <- win_dem %>%
  summarise(SUM= sum(winners))
winners_incumbs1

winners_incumbs2 <- win_rep %>%
  summarise(SUM = sum(winners))
winners_incumbs2

winners_incumbs3 <- function(party_arg) {
  house_elections %>%
  filter(incumbent == "TRUE" &  party == party_arg & ge_winner == "W") %>%
  group_by(state) %>%
  select(state, incumbent, ge_winner, party)
}

lapply(parties, winners_incumbs3)
```

#Percentage of winners. 
Calculating percentage of incumbent democratic winners to total democratic candidates. 
```{r}
percentage_dem <- total_dem %>%
  left_join(win_dem, by = "state") %>%
  mutate(percentage = winners*100/N) %>%
  select(state, percentage)

percentage_dem[is.na(percentage_dem)] <- 0
  
percentage_dem
```

Calculating percentage of incumbent republican winners to total republican candidates
```{r}
percentage_rep <- total_rep %>%
  left_join(win_rep, by = "state") %>%
  mutate(percentage = winners*100/N) %>%
  select(state, percentage)

percentage_rep[is.na(percentage_rep)] <- 0
  
percentage_rep
```
Combining republican percentage data to democratic data by state.
```{r}
percentage_parties <- percentage_dem %>%
  inner_join(percentage_rep, by = "state") %>%
  rename(Dem = percentage.x,
       Rep = percentage.y) %>%
  select(state, Dem, Rep)

list2env(percentage_parties, envir = .GlobalEnv)

percentage_parties
```

**Incumbent Candidates by Party: Why does the candidate I hate keep getting reelected?**

When you have a politician from the opposite party representing your state, it is easy to get annoyed when they continue to get reelected election after election. But do they actually get reelected more often than their peers from the opposite political party? Using data from the Federal Election Commission we explored whether or not incumbent candidates were more likely to get reelected in state elections based on which party they were running with.

So are these incumbent candidates getting reelected more often if they belong to one of the two major political parties in the United States? The answer is that it depends but overwhelmingly no. There are some outlier states that overwhelmingly reelect incumbent candidates from the dominant party of the state. Examples include South Dakota where 100% of incumbent Republican candidates and Maine where 100% of incumbent Democratic candidates were reelected. However, these are exceptions not norms. If you look below at Figure 1, which compares the percentage of Democratic and Republican incumbent candidates reelected by state you see that overwhelmingly there is very little variance based on party.


```{r, message=FALSE, echo=FALSE, warning=FALSE}
ggplot(percentage_parties, aes(y = Rep, x = Dem)) + 
  geom_text(label = state, position = "jitter", aes(angle = 45)) + xlab("Democratic candidates") + ylab("Republican candidates") + labs(title = "Percentage of incumbant candidates reelected by political party")
```
Figure 1: A comparison of the percentage of incumbent Democratic and Republican candidates running in states that got reelected


Using bar charts we chunk the data to make viewing it less overwhelming. These bar charts represent the 50 states with there being one data point per a state. By looking at the bar charts we are able to detect the subtle differences between the data, specifically that the Republican party seems to have more states reelecting more than 20% of their incumbent candidates than the Democratic party.

```{r, message=FALSE, echo=FALSE, warning=FALSE}
Dem_states <- percentage_parties %>%
  filter(Dem > 0) %>%
  group_by(Percentage = cut(Dem, breaks = c(0.1, 20, 40, 60, 80, 100))) %>%
  summarise(state_no2 = n()) %>%
  ggplot(aes(x = Percentage, y = state_no2)) +
  geom_bar(stat = "identity", fill = "blue") 
  

Dem_states + labs(title = "Percentage of Democratic incumbant candidates reelected by state") + xlab("Percentage of incumbant candidates reelected") + ylab("Number of states")
```
Figure 2: The distribution of percentages of Democratic incumbent candidates being reelected by state

```{r, message=FALSE, echo=FALSE, warning=FALSE}
Rep_states <- percentage_parties %>%
  filter(Rep > 0) %>%
  group_by(Percentage = cut(Rep, breaks = c(0.1, 20, 40, 60, 80, 100))) %>%
  summarise(state_no2 = n()) %>%
  ggplot(aes(x = Percentage, y = state_no2)) +
  geom_bar(stat = "identity", fill = "red") 
  

Rep_states + labs(title = "Percentage of Republican incumbant candidates reelected by state") + xlab("Percentage of incumbant candidates reelected") + ylab("Number of states")
```
Figure 3: The distribution of percentages of Republican incumbent candidates being reelected by state





























